---
title: "BM_HW2_yc3729"
author: "Yineng Chen"
date: "2019/9/30"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
rm(list=ls())
```

# Question 1
## a)
```{r}
 # x : the number of cases of uveal melanoma
 # n = 8500000 people, p = 5/1000000 prob
n = 8500000
p = 5/1000000
dbinom(30,n,p)
```
the probability that there are exactly 30
cases occur in NYC in 2018 is `0.00934306`

## b)
```{r}
pn = 6.02/1000000
pb = 0.31/1000000
pa = 0.39/1000000

nn = 8500000 * 0.428
nb = 8500000 * 0.243
na = 8500000 * 0.14

#pn30 = dbinom(30,nn,pn)
#pb30 = dbinom(30,nb,pb)
#pa30 = dbinom(30,na,pa)
```
The probability of having exactly 30 cases occur is `r dbinom(30,nn,pn)` in non-Hispanic Whites,`r dbinom(30,nb,pb)` in black, `r dbinom(30,na,pa)` in Asians.

# Question 2
## a)
```{r import data, include = FALSE}
exer_data = read_csv("./data/HW1.csv",col_types = "cicciiiiiiiiddiiiiii") %>% 
  janitor::clean_names() %>% 
  mutate( group = recode(group, "1" = "intervention","0" = "control"),
          gender = recode(gender, "1" = "male","2" = "female"),
          race = recode(race,"1" = "african american","2" = "hispanic", "3" = "african american", "4" = "Caucasian","5" = "other", "6" = "other"))
```

```{r select data,include = FALSE}
con_bmi = exer_data %>% 
  filter(group == "control") %>% 
  select(bmi_pre,bmi_post) 


inter_bmi = exer_data %>% 
  filter(group == "intervention") %>% 
  select(bmi_pre,bmi_post)

```

Perform two_sample t test for the mean of BMI in control group at 6
months follow-up vs baseline. Testing via build-in functions

### two sample t test for control group

#### Test equality of variances via build-in function
H0: The variances are equal at baseline and end-point.

H1: The variances are not equal at baseline and end-point.
```{r F test}
var.test(pull(con_bmi,bmi_pre),pull(con_bmi,bmi_post),alternative = "two.sided")# accept null hypothesis, the variance are equal
```
H0 is accepted, the variance are equal

#### Test equality of means via build-in function
H0: In control group, there is no difference between the mean of bmi at baseline and  6 months follow-up 

H1: In control group, there is a difference between the mean of bmi at baseline and  6 months follow-up 
```{r t test}
con_bmi_ttest = t.test(pull(con_bmi,bmi_pre),pull(con_bmi,bmi_post), var.equal = TRUE, paired = FALSE) 
con_bmi_ttest# accept null hypothesis, true difference in means is equal to 0
```
Therefore, H0 is accepted, there are no differce between the mean of BMI at baseline and the end-point in control group.



**perform two sample t test via formulation:**

Control group BMI baseline: n1= `r nrow(con_bmi)`, var1 = `r var(pull(con_bmi,bmi_pre))`, x1_bar = `r mean(pull(con_bmi,bmi_pre))`

Control group BMI end-point:n2= `r nrow(con_bmi)`, var2 = `r var(pull(con_bmi,bmi_post))`, x2_bar = `r mean(pull(con_bmi,bmi_post))`

### Test equality of variances via formulation
H0 = variances are equal in two samples.

H1 = variances are different in two samples.
```{r}
num = nrow(con_bmi)
mean_c_pre = mean(pull(con_bmi,bmi_pre))
mean_c_post = mean(pull(con_bmi,bmi_post))
var_c_pre = var(pull(con_bmi,bmi_pre))
var_c_post = var(pull(con_bmi,bmi_post))

F_stat = var_c_pre/var_c_post
F_crit1 = qf(.025, df1 = 35, df2 = 35)
F_crit2 = qf(.975, df1 = 35, df2 = 35)

```
F_stat falls between F_crit1 and F_crit2. Therefore, H0 is accepted variances are equal in two samples.

### Test equality of means via formulation
H0 = means are equal in two samples.
H1 = means are different in two samples.
```{r}
std_pooled = sqrt((var_c_pre^2*(num - 1) + var_c_post^2*(num - 1))/(2*num - 2))
t_stats = (mean_c_pre - mean_c_post)/(std_pooled*sqrt((1/num)+(1/num))) #-0.03265471

t_crit = qt(0.975, 2*num - 2)  #1.994437
```
|t_stats|<t_crit.
H0 is accepted, there are no differce between the mean of BMI at baseline and the end-point in control group.

### two sample t test for intervention group
Test via build-in function:

H0: In intervention group, there is no difference between the mean of bmi at baseline and 6 months follow-up

H1: In control group, there is a difference between the mean of bmi at baseline and 6 months follow-up
```{r}
###Test equality of variances via build-in function################
# Ho: The variances are equal
# H1: The variance are not equal
var.test(pull(inter_bmi,bmi_pre),pull(inter_bmi,bmi_post),alternative = "two.sided") # accept null hypothesis, the variance are equal

###Test equality of mean via build-in function######################
inter_bmi_ttest = t.test(pull(inter_bmi,bmi_pre),pull(inter_bmi,bmi_post), var.equal = TRUE, paired = FALSE) 
inter_bmi_ttest# accept null hypothesis, true difference in means is equal to 0
```
Therefore, H0 is accepted, there are no differce between the mean of BMI at baseline and the end-point in intervention group.

